Realization of a Decoherence-Free Subspace with Nitrogen-Vacancy Centers and $^{13}$C Nuclear Spins

### 2016-09-22 swap-nvst-paper/paper.tex
removed obsolete fixme from paper
added more content to paper pertaining to identifying and locating larmor pairs

% shorthand for frequently used commands
\renewcommand{\u}{\uparrow}
\renewcommand{\d}{\downarrow}
\newcommand{\SWAP}{\t{SWAP}}
\newcommand{\cNOT}{\t{cNOT}}
\newcommand{\NV}{\t{NV}}
\newcommand{\AC}{\t{AC}}
\newcommand{\ST}{\t{ST}}
\renewcommand{\S}{\t{S}}
\newcommand{\T}{\t{T}}
\newcommand{\R}{\mathcal R}

@@

Restricting the NV electron spin states to the subspace spanned by
$\set{\ket0,\ket{m_s}}$ ($m_s=\pm1$), the AXY-$n$ pulse sequence
developed in by Casanova et al. in \cite{casanova2015robust} and
control scheme developed by Wang et al. in \cite{wang2016positioning}
allows addressing individual $^{13}$C nuclear spins via Hamiltonians
of the form
\begin{align}
  H_\t{int} = h_j^\t{int}\sigma_\NV^z\v I_j\c\uv n_j,
  \label{H_int}
\end{align}
for a tunable coupling strength $h_\t{int}$ and direction $\uv n_j$,
where $\sigma_\NV^z\equiv\op{m_s}-\op{0}$ and $\v I_j$ is the spin
operator for nuclear spin $j$. The protocol for realizing this
effective Hamiltonian, however, is conditional on a sufficiently
unique effective larmor frequency $\omega_j$, i.e. that for all
$n\ne j$
\begin{align}
  \abs{\omega_n-\omega_j} \gg \abs{f_{k_{DD}}\v A_j^\perp},
\end{align}
where $f_{k_{DD}}$ is the fourier component of the AXY-$n$ protocol on
resonance with $\omega_j$ and $\v A_j^\perp$ ($\v A_j^\parallel$) is
the component of the hyperfine vector $\v A_j$ perpendicular
(parallel) to the effective larmor precession axis $\uv\omega_j$.

It is possible, however, for two nuclear spins (indexed by 1 and 2) to
have identical larmor frequencies $\omega_1=\omega_2\equiv\omega$,
which occurs when both nuclei lie along the edges of a cylinder
centered on the NV center and oriented along the natural NV axis. We
call such nuclei {\it larmor pairs}, and their geometry results in
$A_1^\perp=A_2^\perp\equiv A^\perp$ and
$A_1^\parallel=A_2^\parallel\equiv A^\parallel$ (see Supplementary
Information) \fixme{prove in the SI}.

\fixme{Reference \cite{wang2016positioning} for identifying and
  location nuclei with unique larmor frequencies}

\fixme{Provide derivation for NV coherence when on resonance with the
  larmor frequency of a larmor pair}

\fixme{Provide derivation for NV coherence when on resonance with a
  larmor frequency when applying a control field for both the one- and
  two-nucleus case}

\fixme{Walk through protocol to identify and locate larmor pairs}

@@

When trying to address one spin in a larmor pair, the protocols for
(\ref{H_int}) will simultaneously couple $\sigma_\NV^z$ to both
$\v I_1\c\uv n_1$ and $\v I_2\c\uv n_2$ with identical coupling
strength ($h_\t{int}$). So long as $\uv n_1$ and $\uv n_2$
(equivalently, $\v A_1^\perp$ and $\v A_2^\perp$) are not parallel,
however, it is possible to address only one of these spins at a time
by applying additional control fields. In particular, the application
of a magnetic decoupling field
$\v B_\t{dec}^{\p\omega}\p{t}=\v B_\t{dec}\cos\p{\omega
  t-\phi_\t{dec}}$ during the AXY-$n$ addressing protocol suppresses
components of $\uv n_j$ which are perpendicular to the vector
$\v B_\t{dec}$ projected onto the plane perpendicular to $\v\omega_j$
and rotated about the axis $\uv\omega_j$ by the angle
$\phi_\t{dec}$. One can therefore choose a direction $\v B_\t{dec}$
which completely suppresses NV electron spin coupling to spin 1 (2),
leaving only coupling to spin 2 (1).

@@

\fixme{Implement PRX\cite{reiserer2016robust} methods for initializing
  a qubit in $\ket{\u\d}\pm\ket{\d\u}$}

@@

@@

@@

The Hamiltonian of the NV center system in the presence of a strong,
static magnetic field $\v B$ can be decomposed as
\begin{align}
  H_\t{sys} = H_\NV^\t{GS} + H_\t{hf} + H_\t{nZ}  + H_\t{nn},
\end{align}
where
\begin{align}
  H_\NV^\t{GS} = D\p{\v S\c\uv z}^2 - \gamma_e\v B\c\v S
\end{align}
is the ground-state NV electron spin Hamiltonian with zero-field
splitting $D$, spin-1 operator $\v S$, natural NV axis $\uv z$, and
electron gyromagnetic ratio $\gamma_e$;
\begin{align}
  H_\t{hf} = \sum_j\f{\gamma_e\gamma_j}{4\pi r_j^3}
  \p{\v S\c\v I_j-3\sp{\v S\c\uv r_j}\sp{\v I_j\c\uv r_j}}
\end{align}
is the hyperfine coupling between the NV electron spin (located at the
origin) and nuclei indexed by $j$ located at $\v r_j$;
\begin{align}
  H_\t{nZ} = -\sum_j\gamma_j\v B\c\v I_j
\end{align}
is the Zeeman Hamiltonian for nuclei indexed by $j$ with gyromagnetic
ratio $\gamma_j$ and spin operators $\v I_j$; and
\begin{align}
  H_\t{nn} = \sum_{j<k}\f{\gamma_j\gamma_k}{4\pi r_{jk}^3}
  \p{\v I_j\c\v I_k-3\sp{\v I_j\c\uv r_{jk}}\sp{\v I_k\c\uv r_{jk}}}
\end{align}
is the internuclear coupling with $\v r_{jk}\equiv\v r_j-\v r_k$. We
will hereafter neglect internuclear coupling in all analytical work,
though we will keep $H_\t{nn}$ in any numerical results.

Changing into the rotating frame of $H_\NV^\t{GS}$ affects only the
hyperfine Hamiltonian, which becomes
\begin{align}
  \tilde H_\t{hf} = S_z\sum_j \v A_j\c\v I_j,
\end{align}
where we define the hyperfine field
\begin{align}
  \v A_j = \f{\gamma_e\gamma_j}{4\pi r_j^3}\p{\uv z - 3\sp{\uv
  r_j\c\uv z}\uv r_j}.
\end{align}
The system Hamiltonian is thus
\begin{align}
  \tilde H_\t{sys} = S_z\sum_j \v A_j\c\v I_j
  -\sum_j\gamma_j\v B\c\v I_j.
\end{align}

When restricted to the subspace of $\set{\ket0,\ket{m_s}}$ with with
$m_s\in\set{1,-1}$ for the NV electron spin,
\begin{align}
  S_z = \f{m_s}2\p{\sigma_\NV^z+\mathbf1},
\end{align}
where $\sigma_\NV^z\equiv\op{m_s}-\op{0}$ and
$\mathbf1\equiv\op{m_s}+\op{0}$. Defining the interaction Hamiltonian
\begin{align}
  H_\t{int} = \f12m_s\sigma_\NV^z\sum_j\v A_j\c\v I_j
\end{align}
the effective larmor frequency
\begin{align}
  \v\omega_j = \gamma_j\v B - \f{m_s}2\v A_j,
\end{align}
and the effective Zeeman Hamiltonian
\begin{align}
  H_\t{nZ}^\t{eff} = -\sum_j\p{\gamma_j\v B - \f{m_s}2\v A_j}\c\v I_j
  = -\sum_j\v\omega_j\c\v I_j,
\end{align}
the system Hamiltonian can be written as
\begin{align}
  \tilde H_\t{sys} = H_\t{int} + H_\t{nZ}^\t{eff}.
  \label{S_H_sys_int_nZ}
\end{align}

Moving into the rotating frame of $H_\t{nZ}^\t{eff}$, the entire
system is described by the Hamiltonian
\begin{align}
  \tilde H_\t{int} = \f12m_s\sigma_\NV^z
  \sum_j\tilde{\v A_j}\p{t}\c\v I_j,
  \label{S_H_int_no_DD}
\end{align}
where $\tilde{\v A_j}\p{t}$ is the vector $\v A_j$ rotated about
$\uv\omega_j$ by an angle $\omega_jt$. Written explicitly,
\begin{align}
  \tilde{\v A_j}\p{t} = \v A_j^\parallel + \v A_j^\perp\cos\p{\omega_jt}
  + \v A_j^{\perp\perp}\sin\p{\omega_jt},
  \label{S_A_rot}
\end{align}
where
\begin{align}
  \v A_j^\parallel\equiv\p{\v A_j\c\uv\omega_j}\uv\omega_j
\end{align}
is the projection of $\v A_j$ along $\v \omega_j$,
\begin{align}
  \v A_j^\perp\equiv\v A_j-\v A_j^\parallel
\end{align}
is the projection of $\v A_j$ onto the plane orthogonal to
$\v\omega_j$, and
\begin{align}
  \v A_j^{\perp\perp}\equiv\uv\omega_j\times\v A_j^\perp
\end{align}
is the vector $\v A_j^\perp$ rotated about $\uv\omega_j$ by $\pi/2$.

\subsection{Dynamic decoupling and nuclear spin addressing}

Dynamic decoupling schemes effectively prepend a modulation function
\begin{align}
  F\p{t}=\sum_kf_k\cos\p{k\omega_{DD}t-\phi_{DD}}
\end{align}
with fundamental frequency $\omega_{DD}$, phase $\phi_{DD}$, and range
$\set{1,-1}$ to $\sigma_\NV^z$, so that the system Hamiltonian reads
\begin{align}
  \tilde H_\t{int} =
  \f12m_sF\p{t}\sigma_\NV^z\sum_j\tilde{\v A_j}\p{t}\c\v I_j.
\end{align}
At a resonance $k_{DD}\omega_{DD}\equiv\omega_r\gg A_j$, the secular
approximation allows one to neglect $\v A_j^\parallel$ and express
\begin{align}
  \tilde H_\t{int} = \f14 f_{k_{DD}} m_s\sigma_\NV^z
  \sum_{j:\omega_j=\omega_r} A_j^\perp\v I_j\c\uv n_j,
  \label{S_H_int_resonance}
\end{align}
where $\uv n_j$ is a unit vector in the direction of $\v A_j^\perp$
rotated about $\uv\omega_j$ by $\phi_{DD}$, i.e.
\begin{align}
  \uv n_j = \uv A_j^\perp\cos\phi_{DD} +
  \uv A_j^{\perp\perp}\sin\phi_{DD}.
\end{align}
One can concisely write (\ref{S_H_int_resonance}) with
\begin{align}
  h_j^\t{int}\equiv f_{k_{DD}}m_sA_j^\perp/4
\end{align}
as
\begin{align}
  \tilde H_\t{int} = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}\sigma_\NV^z\v I_j\c\uv n_j.
  \label{S_H_int}
\end{align}

When the harmonics of $F\p{t}$ are off resonance with all $\omega_j$
and $\omega_{DD}\gg A_j$ for all $j$, the secular approximation allows
one to neglect $\tilde H_\t{int}$ entirely. In this case, one can
apply an auxiliary magnetic field
$\v B_\t{ctl}^{\p\omega}\p{t}=\v
B_\t{ctl}\cos\p{\omega_\t{ctl}t-\phi_\t{ctl}}$ and shift into the
rotating frame of $H_\t{nZ}^\t{eff}$, much like the frame shift
between (\ref{S_H_sys_int_nZ}) and (\ref{S_H_int_no_DD}), to realize
the Hamiltonian
\begin{align}
  \tilde H_\t{sys}^\t{ctl} =
  -\sum_j\gamma_j\tilde{\v B}_{\t{ctl},j}\p{t}
  \cos\p{\omega_\t{ctl}t-\phi_\t{ctl}}\c\v I_j,
\end{align}
with $\tilde{\v B}_{\t{ctl},j}\p{t}$ defined using $\v B_\t{ctl}$ and
$\v\omega_j$ the same way that $\tilde{\v A_j}\p{t}$ was defined using
$\v A_j$ and $\v\omega_j$ in (\ref{S_A_rot}), i.e. as $\v B_\t{ctl}$
rotated about $\uv\omega_j$ by an angle $\omega_jt$:
\begin{align}
  \tilde{\v B}_{\t{ctl},j}\p{t} = \v B_{\t{ctl},j}^\parallel
  + \v B_{\t{ctl},j}^\perp\cos\p{\omega_jt}
  + \v B_{\t{ctl},j}^{\perp\perp}\sin\p{\omega_jt},
\end{align}
where
\begin{align}
  \v B_{\t{ctl},j}^\parallel
  &\equiv \p{\v B_\t{ctl} \c\uv\omega_j}\uv\omega_j \\
  \v B_{\t{ctl},j}^\perp
  &\equiv \v B_\t{ctl} - \v B_{\t{ctl},j}^\parallel \\
  \v B_{\t{ctl},j}^{\perp\perp}
  &\equiv \uv\omega_j\times\v B_\t{ctl}^\perp.
\end{align}

When $\omega_\t{ctl}\gg\gamma_jB_\t{ctl}$, the secular approximation
allows one to neglect $\v B_{\t{ctl},j}^\parallel$ and express
\begin{align}
  \tilde H_\t{ctl} = -\f12\sum_{j:\omega_j=\omega_\t{ctl}}
  \gamma_jB_{\t{ctl},j}^\perp \v I_j\c\uv m_j,
  \label{S_H_ctl_resonance}
\end{align}
where $\uv m_j$ is the direction of $\v B_{\t{ctl},j}^\perp$ rotated
about $\uv\omega_j$ by $\phi_\t{ctl}$, i.e.
\begin{align}
  \uv m_j = \uv B_{\t{ctl},j}^\perp\cos\phi_\t{ctl} +
  \uv B_{\t{ctl},j}^{\perp\perp}\sin\phi_\t{ctl}.
\end{align}
One can concisely write (\ref{S_H_ctl_resonance}) with
\begin{align}
  h_j^\t{ctl}\equiv-\gamma_jB_{\t{ctl},j}^\perp/2
\end{align}
as
\begin{align}
  \tilde H_\t{ctl} = \sum_{j:\omega_j=\omega_\t{ctl}}
  h_j^\t{ctl}\v I_j\c\uv m_j.
  \label{S_H_ctl}
\end{align}

In both (\ref{S_H_int}) and (\ref{S_H_ctl}), one can choose
$\phi_{DD}$ and $\phi_\t{ctl}$ such that $\uv n_j\p\phi$ and
$\uv m_j\p{\phi_\t{ctl}}$ lie anywhere in the plane perpendicular to
$\uv\omega_j$. One can also tune $h_j^\t{int}$ and $h_j^\t{ctl}$
respectively via free parameters in the AXY-$n$ sequence and
$\v B_\t{ctl}$. These controls, together with rotations of the NV
electron spin, allow one to realize a universal set of quantum
gates. An exception to this statement is when two or more nuclei have
the same effective larmor frequency $\omega_j$, in which case they are
addressed simultaneously in (\ref{S_H_int}) and (\ref{S_H_ctl}).

\subsection{Addressing larmor pairs}

We call two spins indexed by $j$ and $k$ elements of a \emph{larmor
  set} if their effective larmor frequencies are equal,
i.e. $\omega_j=\omega_k$, and refer to a larmor set with exactly two
spins as a \emph{larmor pair}. The problem with individually
addressing elements of a larmor set is manifest in (\ref{S_H_int}) and
(\ref{S_H_ctl}), wherein the previously developed procedure for
nuclear spin addressing does not single out any nuclei. Nonetheless,
under certain conditions it is possible to individually address
elements of a larmor set by applying additional control fields.

When employing a dynamic decoupling protocol on resonance with an
effective larmor frequency $\omega_r$, one can apply an additional
magnetic decoupling field
$\v B_\t{dec}^{\p\omega}\p{t}=\v B_\t{dec}\cos\p{\omega_r
  t-\phi_\t{dec}}$ and realize, in the frame of $H_\NV^\t{GS}$, the
system Hamiltonian
\begin{align}
  \tilde H_\t{sys}^\t{dec} = H_\t{int} + H_\t{nZ}^\t{eff} + H_\t{dec},
\end{align}
where
\begin{align}
  H_\t{dec} = -\sum_j\gamma_j\v B_\t{dec}\c\v I_j
  \cos\p{\omega_rt-\phi_\t{dec}}.
\end{align}

>>>>

% shorthand for frequently used commands
\renewcommand{\u}{\uparrow}
\renewcommand{\d}{\downarrow}
\newcommand{\SWAP}{\t{SWAP}}
\newcommand{\cNOT}{\t{cNOT}}
\newcommand{\NV}{\t{NV}}
\newcommand{\AC}{\t{AC}}
\newcommand{\ST}{\t{ST}}
\renewcommand{\S}{\t{S}}
\newcommand{\T}{\t{T}}
\newcommand{\R}{\mathcal R}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}

@@

Restricting the NV electron spin states to the span of
$\set{\ket0,\ket{m_s}}$ ($m_s=\pm1$), the AXY-$n$ pulse sequence
developed in by Casanova et al. in \cite{casanova2015robust} and
control scheme developed by Wang et al. in \cite{wang2016positioning}
allows addressing individual $^{13}$C nuclear spins via Hamiltonians
of the form
\begin{align}
  H_{\t{int}} = h_j^{\t{int}}\sigma_\NV^z\v I_j\c\uv n_j,
  \label{eq:H_int}
\end{align}
for a tunable coupling strength $h_j^{\t{int}}$ and direction
$\uv n_j$, where $\sigma_\NV^z\equiv\op{m_s}-\op{0}$ and $\v I_j$ is
the spin operator for nuclear spin $j$. The protocol for realizing
this effective Hamiltonian is conditional on a sufficiently unique
effective larmor frequency $\omega_j$, i.e. that for all $n\ne j$
\begin{align}
  \abs{\omega_n-\omega_j} \gg \abs{f_{k_{DD}}\v A_j^\perp},
\end{align}
where $f_{k_{DD}}$ is the fourier component of the AXY-$n$ protocol on
resonance with $\omega_j$ and $\v A_j^\perp$ ($\v A_j^\parallel$) is
the component of the hyperfine vector $\v A_j$ perpendicular
(parallel) to the effective larmor precession axis $\uv\omega_j$.

It is possible, however, for two nuclear spins (indexed by 1 and 2) to
have identical larmor frequencies $\omega_1=\omega_2\equiv\omega$,
which occurs when both nuclei lie along the edges of a cylinder
centered on the NV center and oriented along the natural NV axis. We
call such nuclei {\it larmor pairs}, and their geometry implies
$A_1^\perp=A_2^\perp\equiv A^\perp$ and
$A_1^\parallel=A_2^\parallel\equiv A^\parallel$ (see Supplementary
Information) \fixme{prove in the SI}. We wish to identify larmor
pairs and individually address the nuclei making up the pair.

The AXY-$n$ sequence and the Hamiltonian in Eq. (\ref{eq:H_int})
allows for measurement of distinct frequencies $\omega_j$, hyperfine
couplings $A_j^\perp$, and ultimately the positions of $^{13}$C nuclei
via a coupling protocol ending in an NV coherence measurement,
outlined in \cite{wang2016positioning}. When on resonance with several
frequencies $\set{\omega_j}$ which the protocol cannot resolve, the NV
coherence after a time $t$ reads
\begin{align}
  \label{eq:L}
  L\p{t} = \prod_j\cos\p{\f14 f_{k_{DD}}A_j^\perp t}.
\end{align}
After selecting an effective larmor frequency $\omega$ of interest via
the scanning protocol in \cite{wang2016positioning}, coherence
measurements as per Eq. (\ref{eq:L}) for different fourier components
$f_{k_{DD}}$ (or times $t$) are sufficient to determine the number of
nuclei with effective larmor frequency $\omega$. The protocol for
measuring $L$ can thus be used to identify a larmor pair.

After identifying a larmor pair, in order to locate the individual
nuclei in the pair we introduce an auxiliary magnetic decoupling field
$\tilde{\v B}_{\t{dec}}\p{t} = \v
B_{\t{dec}}\cos\p{\omega_{\t{dec}}t-\phi_{\t{dec}}}$. When
$\omega_{\t{dec}}$ is on resonance with the effective larmor frequency
$\omega$, the coherence in Eq. (\ref{eq:L}) is modified as
\begin{align}
  \label{eq:L_dec}
  \tilde L\p{t}
  = \prod_j\cos\p{\f14 f_{k_{DD}}A_j^\perp t\cos\eta_j},
\end{align}
where the angle $\eta_j$ is determined by the directions
$\uv A_j^\perp$ and $\uv B_{\t{dec}}$, as well as the phases
$\phi_{DD}$ and $\phi_{\t{dec}}$ (see Supplementary Info
\ref{sec:addressing_pairs}). Measurements of the coherence $\tilde L$
as a function of fourier components $f_{k_{DD}}$ (or times $t$) and
phases $\phi_{DD}$ thus provides information about the directions
$\uv A_j^\perp$.

\fixme{How do we locate larmor pairs using Eq. (\ref{eq:L_dec})?}

@@

When trying to address one spin in a larmor pair, the protocols for
Eq. (\ref{eq:H_int}) will simultaneously couple $\sigma_\NV^z$ to both
$\v I_1\c\uv n_1$ and $\v I_2\c\uv n_2$ with identical coupling
strength ($h_j^{\t{int}}$). So long as $\uv n_1$ and $\uv n_2$
(equivalently, $\v A_1^\perp$ and $\v A_2^\perp$) are not parallel,
however, it is possible to address only one of these spins at a time
by applying additional control fields. In particular, the application
of a magnetic decoupling field
$\tilde{\v B}_{\t{dec}}\p{t}=\v B_{\t{dec}}\cos\p{\omega
  t-\phi_{\t{dec}}}$ during the AXY-$n$ addressing protocol suppresses
components of $\uv n_j$ which are perpendicular to the vector
$\v B_{\t{dec}}$ projected onto the plane perpendicular to
$\v\omega_j$ and rotated about the axis $\uv\omega_j$ by the angle
$\phi_{\t{dec}}$. One can therefore choose a direction
$\v B_{\t{dec}}$ which completely suppresses NV electron spin coupling
to spin 1 (2), leaving only coupling to spin 2 (1).

@@

@@

\fixme{Provide result for initialization fidelities.}

@@

The protocols developed for addressing larmor pairs can be used to
perform a coherent SWAP operation between the NV center spin and the
DFS qubit. In principle, such a coherent SWAP allows using larmor
pairs as quantum memory registers.

\fixme{Expand.}

@@

The Hamiltonian of the NV center system in the presence of a strong,
static magnetic field $\v B$ can be decomposed as
\begin{align}
  H_{\t{sys}} = H_\NV^{\t{GS}} + H_{\t{hf}} + H_{\t{nZ}}  + H_{\t{nn}},
\end{align}
where
\begin{align}
  H_\NV^{\t{GS}} = D\p{\v S\c\uv z}^2 - \gamma_e\v B\c\v S
\end{align}
is the ground-state NV electron spin Hamiltonian with zero-field
splitting $D$, spin-1 operator $\v S$, natural NV axis $\uv z$, and
electron gyromagnetic ratio $\gamma_e$;
\begin{align}
  H_{\t{hf}} = \sum_j\f{\gamma_e\gamma_j}{4\pi r_j^3}
  \p{\v S\c\v I_j-3\sp{\v S\c\uv r_j}\sp{\v I_j\c\uv r_j}}
\end{align}
is the hyperfine coupling between the NV electron spin (located at the
origin) and nuclei indexed by $j$ located at $\v r_j$;
\begin{align}
  H_{\t{nZ}} = -\sum_j\gamma_j\v B\c\v I_j
\end{align}
is the Zeeman Hamiltonian for nuclei indexed by $j$ with gyromagnetic
ratio $\gamma_j$ and spin operators $\v I_j$; and
\begin{align}
  H_{\t{nn}} = \sum_{j<k}\f{\gamma_j\gamma_k}{4\pi r_{jk}^3}
  \p{\v I_j\c\v I_k-3\sp{\v I_j\c\uv r_{jk}}\sp{\v I_k\c\uv r_{jk}}}
\end{align}
is the internuclear coupling with $\v r_{jk}\equiv\v r_j-\v r_k$. We
will hereafter neglect internuclear coupling in all analytical work,
though we will keep $H_{\t{nn}}$ in any numerical results.

Changing into the rotating frame of $H_\NV^{\t{GS}}$ affects only the
hyperfine Hamiltonian, which becomes
\begin{align}
  \tilde H_{\t{hf}} = S_z\sum_j \v A_j\c\v I_j,
\end{align}
where we define the hyperfine field
\begin{align}
  \v A_j = \f{\gamma_e\gamma_j}{4\pi r_j^3}\p{\uv z - 3\sp{\uv
  r_j\c\uv z}\uv r_j}.
\end{align}
The system Hamiltonian is thus
\begin{align}
  \tilde H_{\t{sys}} = S_z\sum_j \v A_j\c\v I_j
  -\sum_j\gamma_j\v B\c\v I_j.
\end{align}

When NV electron spin is restricted to the span of
$\set{\ket0,\ket{m_s}}$ with $m_s\in\set{1,-1}$,
\begin{align}
  S_z = \f{m_s}2\p{\sigma_\NV^z+\mathbf1},
\end{align}
where $\sigma_\NV^z\equiv\op{m_s}-\op{0}$ and
$\mathbf1\equiv\op{m_s}+\op{0}$. Defining the interaction Hamiltonian
\begin{align}
  H_{\t{int}} = \f12m_s\sigma_\NV^z\sum_j\v A_j\c\v I_j
\end{align}
the effective larmor frequency
\begin{align}
  \v\omega_j = \gamma_j\v B - \f{m_s}2\v A_j,
\end{align}
and the effective Zeeman Hamiltonian
\begin{align}
  H_{\t{nZ}}^{\t{eff}} = -\sum_j\p{\gamma_j\v B - \f{m_s}2\v A_j}\c\v I_j
  = -\sum_j\v\omega_j\c\v I_j,
\end{align}
the system Hamiltonian can be written as
\begin{align}
  \tilde H_{\t{sys}} = H_{\t{int}} + H_{\t{nZ}}^{\t{eff}}.
  \label{eq:S_H_sys_int_nZ}
\end{align}

Moving into the rotating frame of $H_{\t{nZ}}^{\t{eff}}$, the entire
system is described by the Hamiltonian
\begin{align}
  \tilde H_{\t{int}} = \f12m_s\sigma_\NV^z
  \sum_j\v\A_j\p{t}\c\v I_j,
  \label{eq:S_H_int_no_DD}
\end{align}
where $\v\A_j\p{t}$ is the vector $\v A_j$ rotated about $\uv\omega_j$
by an angle $\omega_jt$. Written explicitly,
\begin{align}
  \v\A_j\p{t} = \v A_j^\parallel + \v A_j^\perp\cos\p{\omega_jt}
  + \v A_j^{\perp\perp}\sin\p{\omega_jt},
  \label{eq:S_A_rot}
\end{align}
where
\begin{align}
  \v A_j^\parallel\equiv\p{\v A_j\c\uv\omega_j}\uv\omega_j
\end{align}
is the projection of $\v A_j$ along $\v \omega_j$,
\begin{align}
  \v A_j^\perp\equiv\v A_j-\v A_j^\parallel
\end{align}
is the projection of $\v A_j$ onto the plane orthogonal to
$\v\omega_j$, and
\begin{align}
  \v A_j^{\perp\perp}\equiv\uv\omega_j\times\v A_j^\perp
\end{align}
is the vector $\v A_j^\perp$ rotated about $\uv\omega_j$ by $\pi/2$.

\subsection{Dynamic decoupling and nuclear spin addressing}
\label{sec:dynamic_decoupling}

Dynamic decoupling schemes effectively prepend a modulation function
\begin{align}
  F\p{t}=\sum_kf_k\cos\p{k\omega_{DD}t-\phi_{DD}}
\end{align}
with fundamental frequency $\omega_{DD}$, phase $\phi_{DD}$, and range
$\set{1,-1}$ to $\sigma_\NV^z$, so that the system Hamiltonian reads
\begin{align}
  \tilde H_{\t{int}} =
  \f12m_sF\p{t}\sigma_\NV^z\sum_j\v\A_j\p{t}\c\v I_j.
\end{align}
At a resonance $k_{DD}\omega_{DD}\equiv\omega_r\gg A_j$, the secular
approximation allows one to neglect $\v A_j^\parallel$ and express
\begin{align}
  \tilde H_{\t{int}} = \f14 f_{k_{DD}} m_s\sigma_\NV^z
  \sum_{j:\omega_j=\omega_r} A_j^\perp\v I_j\c\uv n_j,
  \label{eq:S_H_int_resonance}
\end{align}
where $\uv n_j$ is a unit vector in the direction of $\v A_j^\perp$
rotated about $\uv\omega_j$ by $\phi_{DD}$, i.e.
\begin{align}
  \uv n_j = \uv A_j^\perp\cos\phi_{DD} +
  \uv A_j^{\perp\perp}\sin\phi_{DD}.
\end{align}
One can concisely write Eq. (\ref{eq:S_H_int_resonance}) with
\begin{align}
  h_j^{\t{int}}\equiv f_{k_{DD}}m_sA_j^\perp/4
\end{align}
as
\begin{align}
  \tilde H_{\t{int}} = \sum_{j:\omega_j=\omega_r}
  h_j^{\t{int}}\sigma_\NV^z\v I_j\c\uv n_j.
  \label{eq:S_H_int}
\end{align}

When the harmonics of $F\p{t}$ are off resonance with all $\omega_j$
and $\omega_{DD}\gg A_j$ for all $j$, the secular approximation allows
one to neglect $\tilde H_{\t{int}}$ entirely. In this case, one can
apply an auxiliary magnetic field
$\tilde{\v B}_{\t{ctl}}\p{t}=\v
B_{\t{ctl}}\cos\p{\omega_{\t{ctl}}t-\phi_{\t{ctl}}}$ and shift into
the rotating frame of $H_{\t{nZ}}^{\t{eff}}$, much like the frame
shift between Eqs. (\ref{eq:S_H_sys_int_nZ}) and
(\ref{eq:S_H_int_no_DD}), to realize the Hamiltonian
\begin{align}
  \tilde H_{\t{sys}}^{\t{ctl}} =
  -\sum_j\gamma_j\v\B_{\t{ctl},j}\p{t}
  \cos\p{\omega_{\t{ctl}}t-\phi_{\t{ctl}}}\c\v I_j,
\end{align}
with $\v\B_{\t{ctl},j}\p{t}$ defined using $\v B_{\t{ctl}}$ and
$\v\omega_j$ the same way that $\v\A_j\p{t}$ was defined using
$\v A_j$ and $\v\omega_j$ in Eq. (\ref{eq:S_A_rot}), i.e. as
$\v B_{\t{ctl}}$ rotated about $\uv\omega_j$ by an angle $\omega_jt$:
\begin{align}
  \v\B_{\t{ctl},j}\p{t} = \v B_{\t{ctl},j}^\parallel
  + \v B_{\t{ctl},j}^\perp\cos\p{\omega_jt}
  + \v B_{\t{ctl},j}^{\perp\perp}\sin\p{\omega_jt},
\end{align}
where
\begin{align}
  \v B_{\t{ctl},j}^\parallel
  &\equiv \p{\v B_{\t{ctl}} \c\uv\omega_j}\uv\omega_j \\
  \v B_{\t{ctl},j}^\perp
  &\equiv \v B_{\t{ctl}} - \v B_{\t{ctl},j}^\parallel \\
  \v B_{\t{ctl},j}^{\perp\perp}
  &\equiv \uv\omega_j\times\v B_{\t{ctl}}^\perp.
\end{align}

When $\omega_{\t{ctl}}\gg\gamma_jB_{\t{ctl}}$, the secular
approximation allows one to neglect $\v B_{\t{ctl},j}^\parallel$ and
express
\begin{align}
  \tilde H_{\t{ctl}} = -\f12\sum_{j:\omega_j=\omega_{\t{ctl}}}
  \gamma_jB_{\t{ctl},j}^\perp \v I_j\c\uv m_j,
  \label{eq:S_H_ctl_resonance}
\end{align}
where $\uv m_j$ is the direction of $\v B_{\t{ctl},j}^\perp$ rotated
about $\uv\omega_j$ by $\phi_{\t{ctl}}$, i.e.
\begin{align}
  \uv m_j = \uv B_{\t{ctl},j}^\perp\cos\phi_{\t{ctl}} +
  \uv B_{\t{ctl},j}^{\perp\perp}\sin\phi_{\t{ctl}}.
\end{align}
One can concisely write Eq. (\ref{eq:S_H_ctl_resonance}) with
\begin{align}
  h_j^{\t{ctl}}\equiv-\gamma_jB_{\t{ctl},j}^\perp/2
\end{align}
as
\begin{align}
  \tilde H_{\t{ctl}} = \sum_{j:\omega_j=\omega_{\t{ctl}}}
  h_j^{\t{ctl}}\v I_j\c\uv m_j.
  \label{eq:S_H_ctl}
\end{align}

In both Eqs. (\ref{eq:S_H_int}) and (\ref{eq:S_H_ctl}), one can choose
$\phi_{DD}$ and $\phi_{\t{ctl}}$ such that $\uv n_j\p\phi$ and
$\uv m_j\p{\phi_{\t{ctl}}}$ lie anywhere in the plane perpendicular to
$\uv\omega_j$. One can also tune $h_j^{\t{int}}$ and $h_j^{\t{ctl}}$
respectively via free parameters in the AXY-$n$ sequence and
$\v B_{\t{ctl}}$. These controls, together with rotations of the NV
electron spin, allow one to realize a universal set of quantum
gates. An exception to this statement is when two or more nuclei have
the same effective larmor frequency $\omega_j$, in which case they are
addressed simultaneously in Eqs. (\ref{eq:S_H_int}) and
(\ref{eq:S_H_ctl}).

\subsection{Addressing larmor pairs}
\label{sec:addressing_pairs}

We call two spins indexed by $j$ and $k$ elements of a \emph{larmor
  set} if their effective larmor frequencies are equal,
i.e. $\omega_j=\omega_k$, and refer to a larmor set with exactly two
spins as a \emph{larmor pair}. The problem with individually
addressing elements of a larmor set is manifest in
Eqs. (\ref{eq:S_H_int}) and (\ref{eq:S_H_ctl}), wherein the previously
developed procedure for nuclear spin addressing does not single out
any nuclei. Nonetheless, under certain conditions it is possible to
individually address elements of a larmor set by applying additional
control fields.

When employing a dynamic decoupling protocol on resonance with an
effective larmor frequency $\omega_r$, one can apply an additional
magnetic decoupling field
$\tilde{\v B}_{\t{dec}}\p{t}=\v B_{\t{dec}}\cos\p{\omega_r
  t-\phi_{\t{dec}}}$ and realize, in the frame of $H_\NV^{\t{GS}}$,
the system Hamiltonian
\begin{align}
  \tilde H_{\t{sys}}^{\t{dec}}
  = H_{\t{int}} + H_{\t{nZ}}^{\t{eff}} + H_{\t{dec}},
\end{align}
where
\begin{align}
  H_{\t{dec}} = -\sum_j\gamma_j\v B_{\t{dec}}\c\v I_j
  \cos\p{\omega_rt-\phi_{\t{dec}}}.
\end{align}
